---
title: Introduction to sbcdata
output:
    rmarkdown::html_vignette:
        toc_float: true
vignette: >
    %\VignetteIndexEntry{Introduction to sbcdata}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r setup, include = FALSE}
Sys.setenv(LANGUAGE = "en")
library("sbcdata")
sbcdata
```

**Authors**: `r paste0(format(eval(parse(text = packageDescription("sbcdata")[["Authors@R"]])), include = c("given", "family", "role")), collapse =", ")`<br />
**Last modified:** `r file.info("sbcdata.Rmd")$mtime`<br />
**Compiled**: `r date()`

# Introduction

The data were collected as part of the [AMPEL](https://ampel.care) project
(Analysis and Reporting System for the Improvement of Patient Safety
through Real-Time Integration of Laboratory Findings,
\url{https://ampel.care}).
They include administration data, sepsis labels based on ICD10 codes
and laboratory diagnostics from patients admitted to the
University Hospital Leipzig between Januar 2014 and December 2019 and to the
University Hospital Greifswald between Januar 2015 and December 2020,
respectively.
Additionally it provides functions to generate a similar
dataset from MIMIC-IV [@Johnson2021].

## Installation

You can install the released version of `sbcdata`
from directly from github:

```{r install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("ampel-leipzig/sbcdata")
```

## License

The datasets are available under the
[CC-BY](https://creativecommons.org/by/4.0/) license.

## Citation

To cite the `sbcdata` package, please use:

```{r}
citation("sbcdata")
```

# The `sbcdata` package

## Overview

The `sbcdata` dataset contains `ncol(sbcdata)` variables for
`nrow(sbcdata)` laboratory diagnostics.

```{r str}
str(sbcdata)
```

## Missing data

```{r missingdatatbl}
centers <- c("Greifswald", "Leipzig")

labs <- sbcdata[, labcodes$Code, with = FALSE]
labs <- as.matrix(labs)
labsna <- is.na(labs)
mode(labsna) <- "numeric"

nna <- rowsum(labsna, group = sbcdata$Center)
nna <- rbind(nna, Total = colSums(nna))
knitr::kable(t(nna), caption = "Missing Data")

pna <- round(nna / c(table(sbcdata$Center), nrow(sbcdata)) * 100, 2)

knitr::kable(t(pna), caption = "Missing Data [%]")
```

```{r missingdataimg, echo = FALSE}
.imagemd <- function(x, center, subset = 1e4,
                     col = c(palette.colors(2L)[2L], "white")) {
    old.par <- par(no.readonly = TRUE)
    on.exit(par(old.par))

    nr <- min(nrow(x), subset)
    rws <- seq_len(nr)
    nc <- ncol(x)
    nms <- colnames(x)

    mai <- par("mai")
    w <- max(strwidth(nms, "inch"), na.rm = TRUE) + 1/16
    if (mai[2L] < w)
        mai[2L] <- mai[4L] + w # taken from dotchart
    par(mai = mai)

    y <- seq_len(nc)
    image(
        rws, y, x[rws, rev(y)], col = col,
        xlim = 0.5 + c(0L, nr), ylim = 0.5 + c(0L, nc),
        axes = FALSE, xlab = "", ylab = ""
    )
    title(main = paste0("Missing Data: ", center), adj = 0L)
    title(sub = paste0("Subset of ", subset, " rows."), adj = 0L)
    title(xlab = "Rows", adj = 1L)
    mtext(
        rev(nms), at = y, adj = 0L, side = 2L, las = 2L,
        line = (w + 0.1) / par("csi"), cex = 0.75
    )
    abline(h = y, col = "#808080", lty = "dotted", lwd = 1L)
}

for (cnt in centers)
    .imagemd(labsna[sbcdata$Center == cnt,], center = cnt, subset = 1e5)
```

## Densities

```{r densities}
.plotdens <- function(l, main,
                      col = viridisLite::cividis(length(l) + 1)[-1L]) {
    old.par <- par(no.readonly = TRUE)
    on.exit(par(old.par))

    d <- lapply(l, density, na.rm = TRUE)
    xlim <- range(vapply(d, function(x)range(x$x), double(2)))
    ylim <- c(0, max(vapply(d, function(x)max(x$y), double(1))))

    plot(
        NA, xlim = xlim, ylim = ylim,
        axes = FALSE, xlab = "", ylab = "", main = ""
    )
    for (i in seq_along(d)) {
        lines(d[[i]], col = col[i])
        polygon(d[[i]], col = sub("FF$", "55", col[i]))
    }
    title(main = main, adj = 0L)
    title(ylab = "Density", adj = 1L)
    legend(
        "topright",
        legend = paste0(
            names(l), ": ",
            "N = ", vapply(d, function(x)x$n, integer(1)), "; ",
            "BW = ", vapply(d, function(x)formatC(x$bw), character(1))
        ),
        col = col, lty = 1, lwd = 1, cex = 0.75, bty = "n"
    )
    axis(1, lwd.ticks = 0L, col = "#808080")
    axis(2, lwd.ticks = 0L, col = "#808080")
}

for (v in colnames(labs))
    .plotdens(split(labs[, v], sbcdata$Center), v)
```

## Boxplot HGB

```{r boxplothgb}
boxplot(HGB ~ Sex + Center, data = sbcdata, col = viridisLite::cividis(2))
```

## Prevalence

```{r prevalence}
dup <- duplicated(sbcdata[, .(Id, Center)])
tbl <- table(sbcdata[!dup, .(Center, Diagnosis)])
addmargins(tbl)
proportions(tbl, margin = 1)
```

# Session Information

```{r sessionInfo}
sessionInfo()
```

# References
